### 2024-11-22 Article/introduction.tex
minima

The developlment of equality graphs (e-graphs) is to be brought back to the eighties, in the context of automatic theorem provers (ATPs). E-Graphs are then an extension of prior structures, in order to represemt in a more compact way equivalence classes of expressions mantaining closure-under-congruence invariant.
One of the aims of this work is to present a more theoretic representation of e-graphs than the one at the state of the art presented in \cite{egg}, using category theory formalism, in order to integrate the existence of such structure in a research line concerning concurrency of rewrinig-based formalisms.
Indeed, graph-like structures turned out to be very advantageous for distributed system respresentation.
An example of this fact could be a graph as the abstraction a set of entities and relations between them within a state, while rewriting steps model computation steps that, modifying the graph result in changes of the state of the system.
A standard approach for this is the DPO (double pushout) approach, due to its flexibility and locality properties.
A presentation of e-graphs as the one in this work allows to move beyond the operational representation via data structures, in favour of a formalization that model the changes of the structure too.
In this setup, adhesivity results permit to make use of all the advantages on DPO rewriting, because of the generality of such framework.
The elegance of such formalism lies on the definition itself of adhesive categories, namely categories where is reqired of the existence of pullbacks and pushouts along (some) monomorphisms, which are additionally required to be Van Kampen (that is, stability of pushouts under pullbacks and stability of pullbacks under combined pushouts and pullbacks \cite{Ehrig_Golas}).
The only downside of such formalism is the difficulty of proofs using the definition. For cerain calasses of caeogries, there exists a charaterization of such property, presented in \cite{castelnovo2022newcriterionmathcalmmathcalnadhesivity}.

>>>>

The developlment of equality graphs (e-graphs) is to be brought back to the eighties, in the context of automatic theorem provers (ATPs). E-Graphs are then an extension of prior structures, in order to represemt in a more compact way equivalence classes of expressions mantaining closure-under-congruence invariant.
One of the aims of this work is to present a more theoretic representation of e-graphs than the one at the state of the art presented in \cite{egg}, using category theory formalism, in order to integrate the existence of such structure in a research line concerning concurrency of rewrinig-based formalisms.
Indeed, graph-like structures turned out to be very advantageous for distributed system respresentation.
An example of this fact could be a graph as the abstraction a set of entities and relations between them within a state, while rewriting steps model computation steps that, modifying the graph result in changes of the state of the system.
A standard approach for this is the DPO (double pushout) approach, due to its flexibility and locality properties.
A presentation of e-graphs as the one in this work allows to move beyond the operational representation via data structures, in favour of a formalization that model the changes of the structure too.
In this setup, adhesivity results permit to make use of all the advantages on DPO rewriting, because of the generality of such framework.
The elegance of such formalism lies on the definition itself of adhesive categories, namely categories where it is guaranteed the existence of pullbacks and pushouts along (some) monomorphisms, which are additionally required to be Van Kampen (that is, stability of pushouts under pullbacks and stability of pullbacks under combined pushouts and pullbacks \cite{Ehrig_Golas}).
The only downside of such formalism is the difficulty of proofs using the definition. For cerain calasses of caeogries, there exists a charaterization of such property, presented in \cite{castelnovo2022newcriterionmathcalmmathcalnadhesivity}.
